---
title: "Snowy Plover Analyses 2017"
author: "Max Tarjan"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  word_document:
    reference_docx: "style.docx"
  fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.keep="none"}
source("snowy_survival_code.R")
```


#Draft Report prepared for the Bay Delta USFWS
#Introduction
This report provides analyses of fledging success and the adult breeding population size for Western Snowy Plovers in Recovery Unit 3 during the 2017 breeding season.

#Methods
##Fledging and juvenile survival rates
We estimated the fledging rate by tracking the survival of chicks banded within 24 hours after hatching. We restricted our analysis to chicks at sites that were visited at least once a week. These sites were monitored closely, so chicks that went missing before fledging (31 days) were presumed dead. We assumed that their death date fell within a week of the last date they were observed, and we randomly selected the exact day during the 7-day period following their last observation. This assumption may lead to an underestimate in fledging success if the detection rate is sufficiently low. 

We used a survival analysis (survival package in R) to estimate fledging rate and survival over the first 31 days after hatching. We also calculated the juvenile survival rate following the approach by Hudges et al. (2014), with juvenile survival rate defined as the combined probability that an individual both fledged and survived its first winter  (i.e. from hatch date to March 1st of the following year).

##Breeding adult population size
We estimated population sizes of breeding adults by applying N-mixture models to repeated counts of untagged adult birds. All adult birds were counted at least once a week at each site during the 2017 breeding season (May and June). We used the maximum count if multiple counts existed in a week for a single pond. We assumed that counts during the peak of the breeding season (May and June) represent breeding individuals and not migrants. Most birds were of unknown sex, so we estimated the population size for both males and females. 

N-mixture models estimate the regional population size from repeated counts at multiple sites. These models assume negligible immigration and emigration between sites. Adult birds move between ponds daily when foraging, so we aggregated counts to the complex level. We collected sufficient repeated counts at four complexes: Eden Landing, Alviso, Ravenswood, and Warm Springs. We assumed that the true population size at any given complex does not exceed two times the maximum count (i.e. the minimum detection probability is at least 0.5). This assumption corresponds with previous reports that detection probability in San Francisco Bay is 0.65 (Hudgens et al. 2014). We implemented the analyses using the "unmarked" package in R.  We report estimated population sizes for each complex and discuss the limitations of this method given the available data and the mobility of the species. 


#Results
##Fledging and juvenile survival rates

Across all sites, the fledging rate was `r round(length(which(surv.dat$duration==31))/nrow(surv.dat), 2)` and the juvenile survival rate, or the proportion of chicks that fledged and survived through their first winter to March 1 of the following year, was `r round(ticker/nrow(surv.dat),2)` (Figure 1).

The survival curves differ across ponds, with initial drops in survival in the first week at E16B, E6B, and E8, and a more gradual decrease in survival in E14. The initial drop in survival at most ponds may be due to lower effort or detection probability compared to E14. If birds are not tracked closely during the season, all the birds that were never seen again would "die" in the first week according to the model, and all the birds that lived past the season would create a horizontal line in survival, as shown in Figure 1.

##Population size
The detection probability across all surveys and sites was `r round(mean(plogis(coef(fm1, type="det"))),2)`. Estimated population sizes for `r as.character(y$group)` were, respectively, `r round(bup(fm1re),0)` adult breeding birds. 

#Discussion
##Survival and recruitment
We report the fledging rate for Snowy Plovers in recovery Unit 3 for the first time using statistical analyses. At sites that were closely monitored within the Eden Landing complex, the overall fledging rate was `r round(length(which(surv.dat$duration==31))/nrow(surv.dat), 2)` across all years. 

Average juvenile survival (i.e. the probability of fledging and surviving to first winter) across all sites was `r round(ticker/nrow(surv.dat),2)`, which is almost identical to the juvenile survival rate reported by Hudgens et al. (2014) for the south bay: 0.188. 

The initial drop in survival at most ponds may be due to lower effort or detection probability compared to E14. At E14, brood surveys were conducted twice per week during most of 2015 and 2016, while at all other ponds, general breeding plover surveys were conducted only once per week.  E14 and E16B, which had the smallest drop compared to the other two ponds, may have had higher detection probability due to their relatively linear shape, lack of complex topography, and smaller size.  E6B, which had the steepest drop during the first week, is the largest of the four ponds, and also contains a number of topographical features that may obscure detection of broods.  In addition, until 2015, the levee separating E6B and E6A was not drivable, making accurate surveying along much of the interior of the pond very difficult. If birds are not tracked closely during the season, all the birds that were never seen again would "die" in the first week according to the model, and all the birds that lived past the season would create a horizontal line in survival, as shown in Figure 1.

Although the difference in the shapes of the survival curves may be due to differences in survey effort or detection, there is some evidence that survivorship may be affected by site or enhancement. For example, E14 shows a change in the shape of the survival curve following oyster shell enhancement (Figure 2). Survival dropped sharply in the first seven days in 2008, but survival remained high in the first week in all years since 2014. We can rule out the effect of sample size by comparing these plots to those for E6B (Figure 3), which has similar sample sizes in some cases but consistently shows a sharp drop in survival in the first week following hatching. These results indicate that the first week of life may be a critical period, following which survival is more likely. Further, enhancement or other habitat differences at E14 seem to buffer chicks against this initial drop in survival, leading to overall higher fleding rates at this site.  Further investigation is required to rule out the role of detection probability differences across ponds in these conclusions.

##Population size
The N-mixture model indicated that detection probability for adults was `r round(mean(plogis(coef(fm1, type="det"))),2)`. Hudgens reported that detection probability for SF Bay was 0.65 (Hudgens et al. 2014). Our estimate of detection probability is quite low in comparison, and it was likely underestimated due to movement of adults between sites. If birds move between complexes, there is inflated variability in counts (violating the assumption that all birds counted are within the breeding population for that site). Inflated variability leads to underestimates in detection probability and overestimates in population size. This bias is a significant limitation in using N-mixture models with the data that are currently available.

Our knowledge of these sites supports the idea that N-mixture models are greatly overestimating latent population sizes. The left-tailed skew of the posterior probability distribution for Eden Landing (Figure 4) indicates that the latent population size is even greater than the maximum we set for the model (two times the maximum count). Given our great ability to detect birds at this site, we find this estimate to be highly unlikely, and have strong reason to believe that the model is giving inaccurate estimates because one key assumption is violated, namely the assumption that immigration and emigration are negligible. Relaxing that assumption would explain the high variability in our counts across weeks, and would allow the estimate of detection probability to increase with a resultant decrease in estimates of the latent population size. At the moment, we have limited resources that would allow us to estimate immigration/emigration rates between sites on a weekly timescale. We strongly suggest that an effort is made to band and track the movements of adults using telemetry, MOTUS, or another similar method. This would allow for these estimates and for a much more accurate depiction of the breeding population size in these critical breeding habitats.

######page break
#Tables and Figures
Table 1. Sample sizes of banded chicks for each site used to estimate fledging and survival rates.
```{r, echo=F}
knitr::kable(surv.n)
```

######page break
Table 2. Repeated counts of adult Snowy Plovers in Recovery Unit 3 for each week during the 2017 breeding season (May and June).
```{r, echo=F}
knitr::kable(uf.n)
```

######page break
```{r, echo=F, fig.width=6.5, fig.height=4.5, dpi=300, fig.cap="Figure 1. Survival to fledging of snowy plover chicks at closely monitored ponds within the Eden Landing complex in Hayward, CA. Dashed lines denote the 95% CI. n indicates the number of banded chicks that were tracked from within one day of hatching to fledging (31 days). f indicates the proportion of chicks that survived to fledging."}

print(fig.surv1)
```

######page break
```{r, echo=F, fig.width=6.5, fig.height=6.5, dpi=300, fig.cap="Figure 2. Survival to fledging of snowy plover chicks at E14 within the Eden Landing complex in Hayward, CA. Dashed lines denote the 95% CI. n indicates the number of banded chicks that were tracked from within one day of hatching to fledging (31 days). Missing years imply insufficient data (fewer than 10 banded chicks). f indicates the proportion of chicks that survived to fledging."}

print(fig.surv2)
```

######page break
```{r, echo=F, fig.width=6.5, fig.height=6.5, dpi=300, fig.cap="Figure 3. Survival to fledging of snowy plover chicks at E8 within the Eden Landing complex in Hayward, CA. Dashed lines denote the 95% CI. n indicates the number of banded chicks that were tracked from within one day of hatching to fledging (31 days). Missing years imply insufficient data (fewer than 10 banded chicks). f indicates the proportion of chicks that survived to fledging."}

print(fig.surv3)
```

######page break
```{r, echo=F, fig.width=10, fig.height=7, fig.cap="Figure 4. Counts and model estimates (posterior probabilities) of adult birds at four complexes in South San Francisco Bay, CA in 2017. Grey bars denote the frequency of count data and blue lines denote posterior probability distributions of population size estimates from an N-mixture model. Note the left-tailed skew on the posterior probability distribution for Eden Landing, indicating that the latent population size is larger than the maximum allowed by the model."}

print(fig.n)
```

######page break
#Works Cited
Hudgens, BH, Eberhart-Phillips L, Stenzel L, Burns C, Colwell M and Page G (2014). Population Viability Analysis of the Western Snowy Plover. Report prepared for the U.S. Fish and Wildlife Service. Arcata, CA.