---
title: "Snowy Plover Analyses 2017"
author: "Max Tarjan"
date: "December 8, 2017"
output: 
  word_document:
  fig_caption: yes
  reference_docx: "S:/Science/Waterbird/Program Folders (Gulls, SNPL, ADPP, etc)/Snowy Plovers/Snowy survival project 2017/Snowy_plover_analyses_2017.docx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.keep="none"}
source("snowy_survival_code.R")
```


#Draft Report prepared for the Bay Delta USFWS
#Introduction
This report provides analyses of fledging success and the adult breeding population size for Western Snowy Plovers in Recovery Unit 3 during the 2017 breeding season.

#Methods
##Fledging and juvenile survival rates
We estimated the fledging rate by tracking the survival of chicks banded within two days after hatching. We restricted our analysis to chicks at sites that were visited at least once a week. These sites were monitored closely, so chicks that went missing before fledging (31 days) were presumed dead. We assumed that their death date fell within a week of the last date they were observed, and we randomly selected the exact day during the 7-day period following their last observation. This assumption may lead to an underestimate in fledging success if the detection rate is sufficiently low. 

We used a survival analysis (survival package in R) to estimate fledging rate and survival over the first 31 days after hatching. We also calculated the juvenile survival rate following the approach by Hudges et al. (2014), with juvenile survival rate defined as the combined probability that an individual both fledged and survived its first winter  (i.e. from hatch date to March 1st of the following year).

##Breeding adult population size
We estimated population sizes of breeding adults by applying N-mixture models to repeated counts of untagged adult birds. All adult birds were counted at least once a week at each site during the 2017 breeding season (May and June). We used the maximum count if multiple counts existed in a week for a single pond. We assumed that counts during the peak of the breeding season (May and June) represent breeding individuals and not migrants.  

N-mixture models estimate the regional population size from repeated counts at multiple sites. These models assume negligible immigration and emigration between sites. Adult birds move between ponds daily when foraging, so we aggregated counts to the complex level. We collected sufficient repeated counts at four complexes: Eden Landing, Alviso, Ravenswood, and Warm Springs. We assumed that the true population size at any given complex does not exceed two times the maximum count (i.e. the maximum detection probability is at least 0.5). We implemented the analyses using the "unmarked" package in R.  We report estimated population sizes for each complex and discuss the limitations of this method given the available data and the mobility of the species. 


#Results
##Fledging and juvenile survival rates

Across all sites, the fledging rate was `r round(length(which(surv.dat$duration==31))/nrow(surv.dat), 2)` and the juvenile survival rate, or the proportion of chicks that fledged and survived through their first winter to March 1 of the following year, was `r round(ticker/nrow(surv.dat),2)`.

Initial drop in survival at E6B may be due to lower effort or detection probability compared to E14. If birds are not tracked closely during the season, all the birds that were never seen again would "die" in the first week according to the model, and all the birds that lived past the season would create a horizontal line in survival, as shown in Figure 1.

##Population size
The detection probability across all surveys and sites was `r round(mean(plogis(coef(fm1, type="det"))),2)`. Estimated population sizes for `r as.character(y$group)` respectively: `r round(bup(fm1re),0)`. 

#Discussion
##Survival and recruitment
We report the fledging rate for Snowy Plovers in recovery Unit 3 for the first time. At site that were closely monitored within the Eden Landing complex, the overall fledging rate was `r round(length(which(surv.dat$duration==31))/nrow(surv.dat), 2)` across all years. 

Average juvenile survival (i.e. the probability of fledging and surviving to first winter) across all sites was `r round(ticker/nrow(surv.dat),2)`, which is very similar to the juvenile survival rate reported by Hudgens et al. (2014) for the south bay: 0.188. 

Hudgens also reported that Snowy Plovers produced 1.60 male chicks per male, adult survival was 0.71-0.60, and detection probability for SF Bay was 0.65 (Hudgens et al. 2014).

If birds move between complexes, this will cause inflated variability in counts (violating the assumption that all birds counted are within the breeding population for that site) and would lead to underestimates in detection probability and overestimates in population size.

#Tables and Figures
Table 1. Sample sizes of banded chicks for each site used to estimate fledging and survival rates.
```{r, echo=F}
knitr::kable(surv.n)
```

Table 2. Repeated counts of adult Snowy Plovers in Recovery Unit 3 for each week during the 2017 breeding season (May and June).
```{r, echo=F}
knitr::kable(uf.n)
```

```{r, echo=F, fig.width=10, fig.height=7, fig.cap="Figure 1. Survival to fledging of snowy plover chicks at closely monitored ponds within the Eden Landing complex in Hayward, CA. Dashed lines denote the 95% CI. n indicates the number of banded chicks that were tracked from within one day of hatching to fledging (31 days)."}

print(fig.surv1)
```


```{r, echo=F, fig.width=6.5, fig.height=3.75, fig.cap="Figure 2. Survival to fledging of snowy plover chicks at E14 within the Eden Landing complex in Hayward, CA. Dashed lines denote the 95% CI. n indicates the number of banded chicks that were tracked from within one day of hatching to fledging (31 days). Missing years imply insufficient data (fewer than 10 banded chicks)."}

print(fig.surv2)
```


```{r, echo=F, fig.width=6.5, fig.height=3.75, fig.cap="Figure 3. Survival to fledging of snowy plover chicks at E8 within the Eden Landing complex in Hayward, CA. Dashed lines denote the 95% CI. n indicates the number of banded chicks that were tracked from within one day of hatching to fledging (31 days). Missing years imply insufficient data (fewer than 10 banded chicks)."}

print(fig.surv3)
```

```{r, echo=F, fig.width=10, fig.height=7, fig.cap="Figure 4. Counts and model estimates (posterior probabilities) of adult birds at four complexes in South San Francisco Bay, CA in 2017. Grey bars denote the frequency of count data and blue lines denote posterior probability distributions of population size estimates from an N-mixture model."}

print(fig.n)
```

#Works Cited
Hudgens, BH, Eberhart-Phillips L, Stenzel L, Burns C, Colwell M and Page G (2014). Population Viability Analysis of the Western Snowy Plover. Report prepared for the U.S. Fish and Wildlife Service. Arcata, CA.